### 2020-04-06 paper1/article.tex
more

\documentclass[a4paper,UKenglish,cleveref, autoref]{lipics-v2019}
\bibliographystyle{plainurl}
\title{The Impact of Class Immutability on NCSS}
\author{Yegor Bugayenko}
  {Huawei Technologies Co., Ltd.}
  {Russian Research Institute (RRI)}{Moscow, Russia}{yegor.bugayenko@huawei.com}
\authorrunning{Yegor Bugayenko}
\ccsdesc[100]{Software and its engineering~Object oriented languages}
\keywords{Object-Oriented Programming, Constructors, Cohesion}
\usepackage[utf8]{inputenc}
\usepackage[inline]{enumitem}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{pgfplots}
\usepackage{verbatimbox}
\usepackage{interval}
\usepackage{hyperref}
\usepackage{minted}
  \setminted{fontsize=\footnotesize}
  \setminted{breaklines}
  \usemintedstyle{bw}
\newcommand{\code}[1]{\texttt{#1}}
\newenvironment{nicetable}
  {\setlength{\parindent}{0em}\medskip\small}
  {\medskip}
\begin{document}
\maketitle

\begin{abstract}
...
\end{abstract}

\section{Introduction}

One of the most important factors negatively affecting the quality
of object-oriented software is the length of the code. Larger classes
cause readability and maintability issues. One of the simplest way
to calculate size of a Java class is via the NCSS metric, which
stands for Non-Commenting Source Statements. Larger classes with bigger
amount of statements demonstrate higher values of NCSS.

Immutability is a property of a class in OOP. A class is immutable if
it's impossible to modify its attributes after instantiation. In Java this
technically means that all attributes have \texttt{final} modifier attached
to them.

This research empirically validates the relationship between immutability
and NCSS.

\section{Related Work}

No similar research has been done so far.

\section{Empirical Results}

A list of Java repositories were retrieved from GitHub via their
public API. The first 50 repositories were taken, which satisfied
the selection criteria:
\begin{enumerate*}[label={\arabic*)}]
\item more than 1000 stars,
\item more than 200 Kb of data,
\item not archived,
\item public.
\end{enumerate*}
The list included popular Java open source products, like
Spring Framework, RxJava, Google Guava, and others.

Files with \texttt{.java} extension were taken from all repositories.
There were 20,500 files found.

NCSS metric was calculated for each Java class.

Immutability was calculated for each Java class. A class was considered
immutable if it didn't have any attributes without \texttt{final} modified.

The Figure~\ref{fig:1} demonstrates the results obtained. The x-axis is
the value of NCSS. The y-axis is the share of immutable classes among
all classes with the given NCSS.

\begin{figure}[h]
  \input{../ncss.tex}
  \caption{Distribution of classes by NCSS}
  \label{fig:1}
\end{figure}

The tendency is visually obvious: larger classes are less often immutable.

\section{Conclusion}

It was empirically confirmed that larger Java classes tend to be immutable
less frequently than smaller ones.

\end{document}