### 2020-04-07 paper/article.tex
bib

% \documentclass[a4paper,UKenglish,cleveref, autoref]{lipics-v2019}
\documentclass[12pt]{article}
\bibliographystyle{plainurl}
\title{The Impact of Immutability on Class Size}
\author{Yegor Bugayenko\\Huawei Technologies Co., Ltd.\\Russian Research Institute (RRI)\\\texttt{yegor.bugayenko@huawei.com}}{}{}
% \authorrunning{Yegor Bugayenko}
% \ccsdesc[100]{Software and its engineering~Object oriented languages}
% \keywords{Object-Oriented Programming, Immutability, NCSS}
\usepackage[utf8]{inputenc}
\usepackage[inline]{enumitem}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{pgfplots}
\usepackage{verbatimbox}
\usepackage{interval}
\usepackage{hyperref}
\usepackage{minted}
  \setminted{fontsize=\footnotesize}
  \setminted{breaklines}
  \usemintedstyle{bw}
\newcommand{\code}[1]{\texttt{#1}}
\newenvironment{nicetable}
  {\setlength{\parindent}{0em}\medskip\small}
  {\medskip}
\input{../total}
\begin{document}
\raggedbottom
\maketitle

\begin{abstract}
To be written later...
\end{abstract}

\section{Introduction}

One of the most important factors negatively affecting the quality
of object-oriented software is the length of the code. Larger classes
cause readability and maintability issues. One of the simplest way
to calculate size of a Java class is via the NCSS metric, which
stands for Non-Commenting Source Statements. Larger classes with bigger
amount of statements demonstrate higher values of NCSS.

Immutability is a property of a class in OOP. A class is immutable if
it's impossible to modify its attributes after instantiation. In Java this
technically means that all attributes have \texttt{final} modifier attached
to them.
Immutable classes greatly simplify programming, program maintenance, and
reasoning about programs. Immutable classes can be freely shared, even between
concurrent threads and with untrusted code, without the need to worry
about modifications, even temporary ones, that could result in
inconsistent states or broken invariants.
Immutability is a recommended coding practice for Java~\cite{bloch2016}.

This research empirically validates the relationship between immutability
and NCSS.

\section{Related Work}

No similar research has been done so far.

\section{Empirical Results}

A list of Java repositories were retrieved from GitHub via their
public API. The first \totalrepos{} repositories were taken, which satisfied
the selection criteria:
\begin{enumerate*}[label={\arabic*)}]
\item more than 1000 stars,
\item more than 200 Kb of data,
\item not archived,
\item public.
\end{enumerate*}
The list included popular Java open source products, like
Spring, RxJava, Guava, MyBatis, Clojure, JUnit, Lombok,
Graal, Selenium, Spark, Mockito, Neo4j, Jenkins, Netty, and others.

Files with \texttt{.java} extension were taken from all repositories.
There were \totaljavafiles{} files found. Classes without a single
non-\texttt{static} attribute were excluded (such as utility classes,
interfaces, or enums).

NCSS metric and immutability were calculated for each Java class.
A class was considered immutable if it didn't have any
attributes without \texttt{final} modifier.

The Figure~\ref{fig:1} demonstrates the results obtained. The x-axis is
the value of NCSS of Java classes. The y-axis is the share of immutable classes among
all classes with the given NCSS. The diameter of the plot on the graph
is related to the amount of classes found for the specific NCSS. Classes
with NCSS larger than 1000 were excluded from the graph for the sake of
visual elegancy.

\begin{figure}[h]
  \input{../ncss.tex}
  \caption{Distribution of class immutability by NCSS}
  \label{fig:1}
\end{figure}

The tendency is visually obvious: larger classes are less often immutable. It
is also obvious that the majority of classes have NCSS values smaller than 100.

\section{Discussion}

The notion of immutability is not as straightforward as it might seem, and many
different definitions of immutability exist~\cite{potanin2013}.

\section{Conclusion}

It was empirically confirmed that larger Java classes tend to be immutable
less frequently than smaller ones.

The source code of Ruby and Python scripts used to do the research
is available in GitHub repository \texttt{yegor256/size-vs-immutability}.

\bibliography{main}

\end{document}