### 2020-04-09 paper/ncss.tex
summaries

% Max: 3311
% Lo: 2
% Hi: 1000
% Step: 9.98 (steps: 100)
\begin{tikzpicture}
\begin{axis}[width=12cm,height=8cm,
axis lines=middle, xlabel={NCSS}, ylabel={},
xmin=2, xmax=1000, ymin=0, ymax=100,
yticklabel={\pgfmathprintnumber{\tick}\%},
extra tick style={major grid style=black},grid=major]
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(2.0,0)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=4pt] coordinates {
(11.98,51.86260280599904)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=7pt] coordinates {
(21.96,40.19230769230769)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=7pt] coordinates {
(31.94,40.29714978775015)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=8pt] coordinates {
(41.92,36.24282694050136)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=7pt] coordinates {
(51.900000000000006,32.57948213700426)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=6pt] coordinates {
(61.88,30.265995686556437)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=6pt] coordinates {
(71.86,28.994082840236686)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=5pt] coordinates {
(81.84,27.139874739039666)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=5pt] coordinates {
(91.82000000000001,25.547445255474454)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=4pt] coordinates {
(101.80000000000001,24.345679012345677)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=4pt] coordinates {
(111.78,24.395604395604394)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=4pt] coordinates {
(121.76,21.356502242152466)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=3pt] coordinates {
(131.74,18.953846153846154)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=3pt] coordinates {
(141.72,19.180887372013654)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=3pt] coordinates {
(151.70000000000002,19.58997722095672)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=3pt] coordinates {
(161.68,16.338880484114977)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=3pt] coordinates {
(171.66,15.89825119236884)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=2pt] coordinates {
(181.64000000000001,15.646258503401361)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=2pt] coordinates {
(191.62,16.25794732061762)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=2pt] coordinates {
(201.60000000000002,17.51412429378531)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=2pt] coordinates {
(211.58,16.151545363908276)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=2pt] coordinates {
(221.56,13.934426229508196)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=2pt] coordinates {
(231.54000000000002,14.146868250539956)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=2pt] coordinates {
(241.52,14.387633769322235)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(251.5,15.590742996345918)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(261.48,15.031847133757962)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(271.46000000000004,14.701803051317613)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(281.44,12.949640287769784)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(291.42,12.669683257918551)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(301.40000000000003,9.923664122137405)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(311.38,9.748427672955975)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(321.36,9.317803660565724)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(331.34000000000003,10.35120147874307)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(341.32,9.454545454545455)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(351.3,10.054844606946983)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(361.28000000000003,9.496124031007753)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(371.26,9.421841541755889)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(381.24,10.199556541019955)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(391.22,9.647058823529411)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(401.20000000000005,9.156626506024097)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(411.18,9.135802469135802)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(421.16,7.567567567567568)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(431.14000000000004,8.743169398907105)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(441.12,9.63855421686747)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(451.1,9.433962264150944)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(461.08000000000004,9.67741935483871)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(471.06,9.0)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(481.04,8.411214953271028)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(491.02000000000004,7.006369426751593)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(501.0,6.666666666666667)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(510.98,7.241379310344828)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(520.96,6.474820143884892)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(530.94,6.172839506172839)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(540.9200000000001,7.456140350877193)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(550.9,8.071748878923767)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(560.88,6.422018348623854)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(570.86,9.251101321585903)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(580.84,7.456140350877193)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(590.82,8.81057268722467)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(600.8000000000001,8.695652173913043)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(610.78,9.813084112149532)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(620.76,9.047619047619047)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(630.74,8.653846153846153)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(640.72,10.65989847715736)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(650.7,9.473684210526317)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(660.6800000000001,10.05586592178771)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(670.6600000000001,9.35672514619883)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(680.64,10.828025477707007)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(690.62,10.738255033557047)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(700.6,9.45945945945946)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(710.58,10.0)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(720.5600000000001,8.823529411764707)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(730.5400000000001,6.8181818181818175)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(740.52,5.147058823529411)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(750.5,4.794520547945205)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(760.48,4.081632653061225)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(770.46,3.64963503649635)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(780.44,4.166666666666666)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(790.4200000000001,4.25531914893617)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(800.4000000000001,4.316546762589928)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(810.38,4.477611940298507)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(820.36,5.0)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(830.34,3.8095238095238098)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(840.32,6.0606060606060606)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(850.3000000000001,5.208333333333334)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(860.2800000000001,5.747126436781609)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(870.26,4.761904761904762)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(880.24,3.7037037037037033)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(890.22,3.614457831325301)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(900.2,3.7037037037037033)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(910.1800000000001,3.571428571428571)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(920.1600000000001,1.25)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(930.14,1.4084507042253522)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(940.12,0.0)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(950.1,0.0)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(960.08,0.0)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(970.0600000000001,0.0)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(980.0400000000001,0.0)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(990.0200000000001,0.0)};
\end{axis}\end{tikzpicture}